Significant declines in standardised test scores due to COVID-19 school closures disproportionately affect vulnerable students

### 2024-02-15 paper/paper.qmd
Amend the discussion and add in a subtitle

As noted by Jack and Oster [-@Jack_Oster_2023], it can be seen that the equity impacts are two-fold. First, schools (and students) of more vulnerable racial backgrounds -- especially, Black and Hispanic -- are less likely to have access to in-person instruction. Second, for such groups, the consequences of a lack of in-person instruction on academic performance are more severe. This may be due to a smaller investment by such schools in effective remote learning programs or due to the challenges disadvantages students face at home, such as less reliable Internet connection or fewer community resources. Accordingly, while COVID-19 school closures imposed learning loss costs on students as a whole, they affected most negatively those who were already most vulnerable. This exacerbates the already large inequality gap in educational outcomes.

Looking at @fig-race alongside @fig-mathscoresrace and @fig-elascoresrace provides insights into the impacts of virtual schooling. @fig-race demonstrated that districts with higher shares of Black and higher shares of Hispanic students spent more days in virtual schooling as opposed to states with lower shares of these two races. Whilst @fig-mathscoresrace and @fig-elascoresrace show that every race as a whole had lower pass rates in Spring 2021, the difference in pass rates amongst the different races has not significantly changed. This could be evidence that higher rates of virtual schooling did not have a negative impact on test scores, which would indicate that the decline in test scores seen in Spring 2021 was a result of other factors. However, it is important to note that, whilst both @fig-race and @fig-mathscoresrace and @fig-elascoresrace classify districts as being “High” or “Low” based on the shares they comprise of each respective race, Figure \ref{fig-race}’s dataset included data on 48 states, whereas @fig-mathscoresrace and @fig-elascoresrace only had data on 21 states. Moreover, @fig-mathscoresrace and @fig-elascoresrace weighted their results by the total district enrollment. This difference in methodology limits the ability to compare the results represented in the figures.

@@

It is also possible that other types of lockdowns --- such as the closure of counselling services or after-school activities --- might have impacted student learning. This considerable difficulty of separating the effect of school closures from other pandemic-related stressors may bias the results and so affect their validity. One way in which we have helped disentangle these factors is to stratify the test assessments data by race, ELL and FRPL shares per district. By means of this "restriction" technique, we could compare test scores between more homogeneous demographic subgroups and better control for racial and demographic differences that may have impacted student learning outcomes.

@@

Lastly, the state-level assessment data is obtained from only eleven states of the United States, for which district-level data is available. This data set may not be representative of the academic performance of American students as a whole for two reasons. First, state-standardised assessments remain controversial, with critics questioning whether they accurate indicate academic ability [@yang]. Certainly, they cannot capture the ways in which students learned that were not directly given in the assessment. Second, the sample of eleven states may be too small to be generalisable. Indeed, Parolin and Kee [-@parolin] showed that wide geographic disparities existed in school closures across the country: those concentrated on the West Coast faced declines of \>75% in-person visits to schools, whereas those in the Midwest faced declines of \<25%. Both of these factors put into question the validity and generalisability of the state-assessment data, and could present a valuable direction for future work.

>>>>

As noted by Jack and Oster [-@Jack_Oster_2023], it can be seen that the equity impacts are two-fold. First, as shown in @fig-race, schools (and students) of more vulnerable racial backgrounds -- especially, Black and Hispanic -- are less likely to have access to in-person instruction. Second, as shown in @fig-mathscoresrace and @fig-elascoresrace, for such groups, the consequences of a lack of in-person instruction on academic performance are more severe.

This may be due to a smaller investment by such schools in effective remote learning programs or due to the challenges disadvantages students face at home, such as less reliable Internet connection or fewer community resources. Accordingly, while COVID-19 school closures imposed learning loss costs on students as a whole, they affected most negatively those who were already most vulnerable. This exacerbates the already large inequality gap in educational outcomes.

@@

It is also possible that other types of lockdowns --- such as the closure of counselling services or after-school activities --- might have impacted student learning. This considerable difficulty of separating the effect of school closures from other pandemic-related stressors may bias the results and so affect their validity. One way in which we have helped disentangle these factors is to stratify the test assessments data by race shares per district. By means of this "restriction" technique, we could compare test scores between more homogeneous demographic subgroups and better control for racial differences that may have impacted student learning outcomes.

@@

Another important limitation is the discrepancy in the state information provided by the datasets and in the methodology used to analyse them. In particular, whilst the datasets to create @fig-race contained data on all 48 states, @fig-mathscoresrace and  @fig-elascoresrace had data on only 21 states. Moreover, the results in @fig-mathscoresrace and @fig-elascoresrace were calculated by weighting the averages by total district enrollment, whereas no such weightage was performed for @fig-race. These such differences limit our ability to compare the results represented in the figures.

Lastly, the state-level assessment data is obtained from 21 states of the United States, for which district-level data is available. This data set may not be representative of the academic performance of American students as a whole for two reasons. First, state-standardised assessments remain controversial, with critics questioning whether they accurate indicate academic ability [@yang]. Certainly, they cannot capture the ways in which students learned that were not directly given in the assessment. Second, Jack and Oster [@Jack_Oster_2023] gave no justification for their choice of these 21 states in particular, and it may be the case that the sample is not representative of the United States as a whole.  Indeed, Parolin and Kee [-@parolin] showed that wide geographic disparities existed in school closures across the country: those concentrated on the West Coast faced declines of \>75% in-person visits to schools, whereas those in the Midwest faced declines of \<25%. It remains to examine whether these 21 states form a representative sample of the wide range of school closures across the country. Both of these factors put into question the validity and generalisability of the state-assessment data, and could present a valuable direction for future work.
